Cluster Difference Imaging Photometric Survey. III. An Open Cluster With a 300\,pc Halo

### 2021-03-05 cluster_paper/ms.tex
update ms

% NB. the "classical tidal radius" is from e.g., Piskunov+08 or
% Kharchenko+2013. The latter givs a radius r2 of 0.71 degrees. At 380
% pc, corresponds to 4.7pc. The King tidal radius given by K13 is
% 7.70pc.
\begin{abstract}
  Recent analyses of the Gaia data have reported the existence of
  diffuse stellar populations surrounding nearby open clusters. 
  The stars in these diffuse halos could be the tidal escapees of the
  cluster cores, or they could be remnants of an initially dispersed
  star formation complex.
  They could also be field stars---the false positives of overly
  aggressive clustering algorithms.
  In this study, we focus on a halo with a length of 300\,pc reported
  around the $\approx$175\,Myr open cluster NGC\,2516.
  The classical tidal radius of the cluster is $\lesssim$10\,pc.
  Combining photometry from Gaia, rotation periods from TESS, and
  lithium measurements from Gaia-ESO and GALAH, we find that the halo
  of NGC\,2516 is real and coeval with the core.
  Two in three kinematically selected halo members 25-250\,pc from the
  cluster center show rotation periods consistent with a
  gyrochronological age of 120\,Myr.
  A comparison sample of field stars shows no such trend.
  The rotation period and lithium abundances of the cluster stars are
  also anti-correlated at fixed stellar mass, as has been observed in
  similar stellar populations.
  This work expands the set of confirmed NGC\,2516 members by a factor
  of $\approx$2, and quantifies the degree to which Gaia-based
  analyses can identify dispersed stellar populations against a
  background of field star contaminants.
  We highlight the implications for spectroscopic survey targeting,
  open cluster dispersal, and planet searches around young stars.
\end{abstract}

@@

In the right panel of Figure~\ref{fig:hr}, we compare the observed
Gaia EDR3 photometry with PARSEC isochrones
\citep{bressan_parsec_2012,chen_improving_2014,chen_parsec_2015,marigo_new_2017}.
We used the web
interface\footnote{\url{http://stev.oapd.inaf.it/cgi-bin/cmd},
\texttt{2021-02-26}, \texttt{CMD3.4}, \texttt{YBC} bolometric
corrections as in \citet{chen_2019_YBC}.} to interpolate these
isochrones at $\log_{10}(t/\mathrm{yr})=\{8, 8.25, 8.5\}$, and applied
the reddening correction described in Section~\ref{subsec:tess}.  We
assumed a reddening of $E(B-V)=0.102\pm0.019$, $R_V=3.1$, and applied
the \citet{odonnell_1994} SED-dependent extinction law star-by-star
through the PARSEC interface.  For the metallicity, we considered a
range of super and sub-solar metallicities to fit as much of the locus
from $0.5<Bp-Rp<1.5$ as possible, and settled on the slightly
super-solar $[M/H]=0.06$.  Sub-solar metallicities led to model
predictions too blue along the main sequence by $\approx$0.1\,mag.
Our adopted metallicity agrees with that found by
\citet[][Sec~4.4.4]{cummings_2011_li_iron}, and with the iron
abundance recently determined by the Gaia-ESO team
\citep{baratella_gaiaeso_2020}, which represented an up-revision from
an earlier sub-solar metallicity \citep{randich_gaiaeso_2018}.  While
\citet{bailey_rv_2018} report a slightly sub-solar metallicity for the
cluster, we prefer super-solar based on the goodness of fit from the
photometry.

Overall, the data and models agree for masses above approximately
0.45\,$M_\odot$.  However below this mass, the data and models diverge
at colors redder than Bp-Rp$\approx2.2$.  The MIST isochrones showed a
comparable disagreement \citep{choi_mesa_2016}.  Explanations invoking
both starspots and incomplete molecular line lists have been proposed
\citep[{\it
e.g.},][]{stauffer_why_2003,feiden_magnetic_2013,rajpurohit_effective_2013,mann_spectrothermometry_2013,choi_mesa_2016}.
We adopt the PARSEC isochrones since they diverge at slightly lower
mass than the MIST models, due to the temperature-opacity calibration
implemented by \citet{chen_improving_2014}.  For purposes of
age-dating the cluster, we prefer to simply focus on the main-sequence
turn-off (MSTO), since it is where the models have maximal predictive
power.

@@

%FIXME: except actually, will need to redo with the new source list
%too .......
First, we collected the \nkinematic Gaia DR2 source\_ids corresponding
to the kinematically-identified core and halo members.  For each
source, we first retrieved all available CDIPS light curves, on a
per-sector basis.  This yielded \ncdips light curves, all brighter
than $Rp=16$\,mag.  The $Rp=16$\,mag cutoff imposed during the CDIPS
processing corresponds roughly to $(Bp-Rp)_0$ of 2.2, or a spectral
type of $\approx$M2V, at the distance of NGC\,2516.

@@

To put the clusters on the same color scale, we needed to adopt
reddening corrections.
We assumed {X} for the Pleiades, and {Y} for Praesepe.
For NGC\,2516, we queried the 2MASS DUST service to find the
extinction parameters for the positions of each NGC\,2516 member
(CITE: astrobase.services.dust; 2MASS DUST).  The mean and standard
deviation of the E(B-V) values for the Schlegel+98 and SF11 maps were
consistent, so we considered those of Schlegel+1998:
$E(B-V)=0.206\pm0.039$.  Bonifacio+2000 noted though that Schlegel+98
overestimate the reddening values when the color excess exceeds about
0.10 mag. We applied their correction (Eq 1 from Bonifacio+2000),
which led to an adopted $E(B-V)=0.102\pm0.019$.  Finally, to convert
to Gaia colors we adopted the calibration of Stassun et al 2019, and
assumed $E(Bp-Rp)=1.31 E(B-V)$ and $A_G=2.72 E(B-V)$.  This yielded
$E(Bp-Rp)=0.134\pm0.025$, which is used in the plots.

@@

>>>>

% NB. the "classical tidal radius" is from e.g., Piskunov+08 or
% Kharchenko+2013. The latter givs a radius r2 of 0.71 degrees. At 380
% pc, corresponds to 4.7pc. The King tidal radius given by K13 is
% 7.70pc.
\begin{abstract}
  Recent analyses of the Gaia data have reported the existence of
  diffuse stellar populations surrounding nearby open clusters. 
  The stars in these diffuse halos could be the tidal escapees of the
  cluster cores, or they could be remnants of an initially dispersed
  star formation complex.
  They could also be field stars---the false positives of overly
  aggressive clustering algorithms.
  In this study, we focus on a halo with a length of 300\,pc reported
  around the $\approx$150\,Myr open cluster NGC\,2516.
  The classical tidal radius of the cluster is $\lesssim$10\,pc.
  Combining photometry from Gaia, rotation periods from TESS, and
  lithium measurements from Gaia-ESO and GALAH, we find that the halo
  of NGC\,2516 is real and coeval with the core.
  Two in three kinematically selected halo members 25-250\,pc from the
  cluster center show rotation periods consistent with a
  gyrochronological age of 120\,Myr.
  A comparison sample of field stars shows no such trend.
  The rotation period and lithium abundances of the cluster stars are
  also anti-correlated at fixed stellar mass, as has been observed in
  similar stellar populations.
  This work expands the set of confirmed NGC\,2516 members by a factor
  of $\approx$2, and quantifies the degree to which Gaia-based
  analyses can identify dispersed stellar populations against a
  background of field star contaminants.
  We highlight the implications for spectroscopic survey targeting,
  open cluster dispersal, and planet searches around young stars.
\end{abstract}

@@

In the right panel of Figure~\ref{fig:hr}, we compare the observed
Gaia EDR3 photometry with PARSEC isochrones
\citep{bressan_parsec_2012,chen_improving_2014,chen_parsec_2015,marigo_new_2017}.
We used the web
interface\footnote{\url{http://stev.oapd.inaf.it/cgi-bin/cmd},
\texttt{2021-02-26}, \texttt{CMD3.4}, \texttt{YBC} bolometric
corrections as in \citet{chen_2019_YBC}.} to interpolate these
isochrones at $\log_{10}(t/\mathrm{yr})=\{8, 8.25, 8.5\}$, and applied
the reddening correction described in Section~\ref{subsec:tess}.  We
assumed a reddening of $E(B-V)=0.102\pm0.019$, $R_V=3.1$, and applied
the \citet{odonnell_1994} SED-dependent extinction law star-by-star
through the PARSEC interface.  For the metallicity, we considered a
range of super and sub-solar metallicities to fit as much of the locus
from $0.5<Bp-Rp<1.5$ as possible, and settled on the slightly
super-solar $[M/H]=0.06$ \citep{cummings_2011_li_iron}.  Sub-solar
metallicities led to model predictions too blue along the main
sequence by $\approx$0.1\,mag.  Our adopted metallicity agrees with
the spectroscopic metallicity from
\citet[][Sec~4.4.4]{cummings_2011_li_iron}, and with the iron
abundance recently determined by the Gaia-ESO team
\citep{baratella_gaiaeso_2020}, which represented an up-revision from
an earlier sub-solar metallicity \citep{randich_gaiaeso_2018}.  While
\citet{bailey_rv_2018} report a slightly sub-solar metallicity for the
cluster, we prefer super-solar based on the goodness of fit from the
photometry.

Overall, the data and models agree for masses above approximately
0.45\,$M_\odot$.  However below this mass, the data and models diverge
at colors redder than Bp-Rp$\approx2.2$, in the sense that the model
isochrones have lower luminosities and bluer colors than the
observations.  The MIST isochrones showed a comparable disagreement
\citep{choi_mesa_2016}.  Explanations invoking both starspots and
incomplete molecular line lists have been proposed \citep[{\it
e.g.},][]{stauffer_why_2003,feiden_magnetic_2013,rajpurohit_effective_2013,mann_spectrothermometry_2013,choi_mesa_2016}.
We adopt the PARSEC isochrones since they diverge at slightly lower
mass than the MIST models, due to the temperature-opacity calibration
implemented by \citet{chen_improving_2014}.  For purposes of
age-dating the cluster, we prefer to focus on the main-sequence
turn-off (MSTO), since this is where the models have maximal predictive
power.

@@

We began our rotation analysis by considering all \nkinematic\
candidate cluster members.  For each source, we first retrieved all
available CDIPS light curves, on a per-sector basis.

%FIXME: except actually, will need to redo with the new source list
%too .......

This yielded \ncdips light curves, all brighter
than $Rp=16$\,mag.  The $Rp=16$\,mag cutoff imposed during the CDIPS
processing corresponds roughly to $(Bp-Rp)_0$ of 2.2, or a spectral
type of $\approx$M2V, at the distance of NGC\,2516.

@@

To put the clusters on the same color scale, we needed to adopt
reddening corrections.
% FIXME: where did these reference tables come from???
We assumed {X} for the Pleiades, and {Y} for Praesepe.
For NGC\,2516, we used the \texttt{astrobase.services.dust} module to
query the 2MASS DUST
service\footnote{\url{http://irsa.ipac.caltech.edu/applications/DUST}}
and retrieve the extinction parameters for the positions of each
NGC\,2516 member \citep{bhatti_astrobase_2018}.  The mean and standard
deviation of the $E(B-V)$ values for the \citet{schlegel_maps_1998}
and \citet{schlafly_measuring_2011} maps were consistent, so we
adopted the average from \citet{schlegel_maps_1998}:
$E(B-V)=0.206\pm0.039$.  \citet{bonifacio_search_2000}
\citet{schlegel_maps_1998} maps overestimate the reddening values when
the color excess exceeds about 0.10 mag. Applying the correction
proposed by \citet{bonifacio_search_2000} leads to our adopted
$E(B-V)=0.102\pm0.019$.  Finally, to convert to Gaia colors we adopted
the calibration of \citet{stassun_TIC8_2019}, namely $E(Bp-Rp)=1.31
E(B-V)$ and $A_G=2.72 E(B-V)$.  This yielded $E(Bp-Rp)=0.134\pm0.025$,
which is used in the plots.

@@

\subsection{Absolute age of NGC 2516}
Cummings & Kalirai (2018) favor 150 Myr based on MSTO.
Godoy-Rivera+2021 favor 150 Myr... based on ...? what?
"""
Similar to Cummings et al. (2016), Cummings & Kalirai (2018),
and Cummings et al. (2018) (see also Fritzewski et al. 2020 and Healy
&
McCullough 2020)
"""
